---
title: "Territory size statistical analysis"
output: html_notebook
author: "Ruth Simberloff"
date: "March 2022"
---

###Set-up
```{r load_tidyverse, echo=FALSE}
library(tidyverse)
library(ggsignif)
library(car)
library(fBasics)
library(ggsignif)
library(AICcmodavg)
```


We used mean communication distance per bird to prevent the analysis from weighting certain birds more heavily just because they have more amplitude measures.

```{r read_data}

summary_data <- read_csv("model_data.csv")

```


Table 1. Summarize the data we collected at each site.
```{r site_summary_stats}

site_summary <- summary_data %>%
  group_by(dialect) %>%
  summarise(n = n(),
            SY = sum(age=="SY"),
            median_noise = median(LAF90),
            range_noise = max(LAF90)-min(LAF90),
            median_comm_dist = median(comm_distance),
            mean_body_con = mean(body_condition),
            median_area = median(area_75))

```



# Statistical analysis

### Linear model

This model includes   
- communication distance (25% discrimination distance)
- mean background noise on territory
- bird age (SY/ASY)  
- body condition  
as possible predictors


``` {r candidate_models}


candidate_lms <- list()

candidate_lms[[1]] <- lm(area_75 ~comm_distance, data = summary_data)

candidate_lms[[2]] <- lm(area_75 ~comm_distance + age, data = summary_data)

candidate_lms[[3]] <-  lm(area_75 ~comm_distance + body_condition, data = summary_data)

candidate_lms[[4]] <-  lm(area_75 ~comm_distance + LAF90, data = summary_data)

candidate_lms[[5]] <-  lm(area_75 ~comm_distance + age + body_condition, data = summary_data)

candidate_lms[[6]] <-  lm(area_75 ~comm_distance + age + LAF90, data = summary_data)

candidate_lms[[7]] <-  lm(area_75 ~comm_distance + body_condition + LAF90, data = summary_data)

candidate_lms[[8]] <-  lm(area_75 ~comm_distance + age + body_condition + LAF90, data = summary_data)

candidate_lms[[9]] <-  lm(area_75 ~age, data = summary_data)

candidate_lms[[10]] <- lm(area_75 ~age + body_condition, data = summary_data)

candidate_lms[[11]] <-  lm(area_75 ~age + LAF90, data = summary_data)

candidate_lms[[12]] <-  lm(area_75 ~age + body_condition + LAF90, data = summary_data)

candidate_lms[[13]] <-  lm(area_75 ~body_condition, data = summary_data)

candidate_lms[[14]] <-  lm(area_75 ~body_condition + LAF90, data = summary_data)

candidate_lms[[15]] <-  lm(area_75 ~LAF90, data = summary_data)

candidate_lms[[16]] <-  lm(area_75 ~ 1, data = summary_data)

candidate_lm_names <- c("communication distance",
              "communication distance + age",
              "communication distance + body condition",
              "communication distance + noise",
              "communicaiton distance + age + body condition",
              "communication distance + age + noise",
              "communication distance + body condition + noise",
              "communication distance + age + body condition + noise",
              "age",
              "age + body condition",
              "age + noise",
              "age + body condition + noise",
              "body condition",
              "body condition + noise",
              "noise",
              "null")
```

AICc model selection
Table A1.
```{r model_selection}

aictab(candidate_lms, candidate_lm_names)

```

Evaluating the top-ranked model
```{r top_ranked_model}

plot(candidate_lms[[2]])
normalTest(resid(candidate_lms[[2]]))

summary(candidate_lms[[2]])


```

### Urban/rural comparisons

Urban birds have significantly shorter communication distance.
```{r UR_comm_distance}

rural_distance <- summary_data %>%
  dplyr::filter(habitat == "Rural") %>%
  dplyr::select(comm_distance)

urban_distance <- summary_data %>%
  dplyr::filter(habitat == "Urban") %>%
  dplyr::select(comm_distance)

t.test(urban_distance, rural_distance)
```

Urban birds have significantly smaller territories
```{r UR_area75}

rural_areas <- summary_data %>%
  dplyr::filter(habitat == "Rural") %>%
  dplyr::select(area_75)

urban_areas <- summary_data %>%
  dplyr::filter(habitat == "Urban") %>%
  dplyr::select(area_75)


## This does a Welch-Satterthwaite two-sample t-test, which compensates for the unequal variances of the two groups.
## Note that I square-root transform the data so that they're normally distributed
t.test(sqrt(rural_areas), sqrt(urban_areas))

## comparing differences in means from untransformed data 
mean(rural_areas$area_75)
sd(rural_areas$area_75) / sqrt(length(rural_areas$area_75)) # standard error

mean(urban_areas$area_75)
sd(urban_areas$area_75) / sqrt(length(urban_areas$area_75)) # standard error

# The mean rural territory is is 5440 square meters, while the mean urban territory size is 2328 square meters -- less than half the size.
```


### Background noise is strongly correlated with communication distance
``` {r noise_comm_distance}

lm(comm_distance ~ LAF90, data = summary_data) %>%
summary()

```



Visualization

Figure 2.
```{r regression_plot}


pred <- predict.lm(lm(area_75~comm_distance,  data = summary_data), 
                   se.fit = TRUE, interval = "confidence")
limits <- as.data.frame(pred$fit)


ggplot( data = summary_data, 
       mapping = aes(x = comm_distance, 
                     y = area_75)) +
   geom_line(aes(x = comm_distance, 
                y = limits$lwr),
            linetype = 2, 
            color = "gray50") +
  geom_line(aes(x = comm_distance, 
                y = limits$upr), 
            linetype = 2, 
            color = "gray65") +
  geom_smooth(method = "lm", 
              color = "black", 
              fill = NA) +
  geom_point(size = 1.7, mapping = aes(shape = age), color = "black", fill = "gray80", stroke = .3) +
  scale_shape_manual(values = c(19, 21)) +
  labs(
    shape = "Age",
      x = "Communication distance", 
       y = expression(paste("Territory area ", (m^2)))) +
  theme_linedraw() +
  coord_cartesian(xlim = c(5, 85))
  
```

Figure 3.
```{r UR_boxplots}

# Figure 3 (a)
ggplot(data = summary_data, 
                              mapping = aes(x = habitat, y = comm_distance, fill = habitat)) +
  geom_boxplot(outlier.shape = NA) +
  scale_fill_manual(values = c("#0990B0", "#E23C33")) +
  geom_point(shape = 21, size = 2.4, stroke = .6, alpha = 0.5) +
  geom_signif(comparisons = list(c("Rural", "Urban")), 
              map_signif_level = TRUE,
              tip_length = 0.01) +
  labs(x = "Habitat type", 
       y = "Communication distance (m)") +
  theme_linedraw()

# Figure 3 (b)
ggplot(data = summary_data, 
                              mapping = aes(x = habitat, y = area_75, fill = habitat)) +
  geom_boxplot(outlier.shape = NA) +
  scale_fill_manual(values = c("#0990B0", "#E23C33")) +
  geom_point(shape = 21, size = 2.4, stroke = .6, alpha = 0.5) +
  geom_signif(comparisons = list(c("Rural", "Urban")), 
              map_signif_level = TRUE,
              tip_length = 0.01) +
  labs(x = "Habitat type", 
       y = expression(paste("Territory area ", (m^2)))) +
  theme_linedraw()


```


Figure A3.
```{r noise_comm_distance_plot}
pred <- predict.lm(lm(comm_distance~LAF90,  data = summary_data), 
                   se.fit = TRUE, interval = "confidence")
limits <- as.data.frame(pred$fit)

ggplot(aes(x = LAF90, y = comm_distance), data = summary_data) +
  geom_line(aes(x = LAF90, 
                y = limits$lwr),
            linetype = 2, 
            color = "gray50") +
  geom_line(aes(x = LAF90, 
                y = limits$upr), 
            linetype = 2, 
            color = "gray65") +
  geom_smooth(method = "lm", 
              color = "black", 
              fill = NA) +
  geom_point(size = 1.7) +
  labs(x = "Background noise (LAF90; dB)", 
       y = "Communication distance (m)") +
  theme_linedraw() 

```
